\framework: Morphology-based Entity and Relational Entity Extraction Framework for Arabic

### 2018-04-20 tex/merf/TALMajor/results.tex
done till 3.19

In this section we compare \framework implementations of 
the narrator chain, temporal entity, and genealogy entity extration 
tasks to the task specific techniques proposed to solve them
in ANGE~\cite{ZaMaFlairs2012HadithBio}, 
ATEEMA~\cite{ZaMa2012IJCLATime},  and
GENTREE~\cite{ZaMaHaCicling2012Entity}, respectively. 
We also compare a \framework number normalization task to 
a task specific implementation. 
In the online appendix%
~\footnote{available at ~\url{-ommited-for-anonymity-}}%http://webfea.fea.aub.edu.lb/fadi/pdfs/merfappendix.pdf
, we report on eight additional \framework case studies.

Table~\ref{tab:results} reports the development time,
extraction runtime, recall and precision 
of the output MRE tags, 
the size of the task in lines of code or in number of \framework rules, 
for both the standalone task specific and the \framework implementations. 
%Table~\ref{tab:mbfer} reports the accuracy of the output MBF tags and the user-defined relation 
%construction in \framework for each task.

% Table generated by Excel2LaTeX from sheet 'Sheet1'
\begin{table}[tb!]
  \centering
\caption{\framework compared to task specific applications.}
  \resizebox{0.85\columnwidth}{!}{
    \begin{tabular}{l|l|l|ll|l}
     \toprule
     \multirow{2}{*}{Task} & Development& Run & \multicolumn{2}{c|}{Accuracy} & \multirow{2}{*}{Ease of Composition}\\
     & time & time(s) & Recall & \multicolumn{1}{c|}{Precision} & \\
    \midrule
    ANGE~\cite{ZaMaFlairs2012HadithBio}
    & 2 months & 1.79 & 0.99 & 0.99 & 3000+ lines of code\\
    \framework & 3 hours & 7.24 & 0.99 & 0.93  & 8 MBFs and 4 MREs\\
    \midrule
    ATEEMA~\cite{ZaMa2012IJCLATime}
    & 1.5 months & 2.53 & 0.88  & 0.89  & 1000+ lines of code  \\
    \framework & 3 hours & 3.14 & 0.91  & 0.81  & 3 MBFs and 2 MREs\\
    \midrule
    Genealogy tree~\cite{ZaMaHaCicling2012Entity} 
    & 3 weeks & 0.74 & 0.96 & 0.98 & 3000+ lines of code\\
    \framework & 4 hours & 2.28 & 0.84 & 0.93  & 3 MBFs and 3 MREs\\
    \midrule
    NUMNORM & 1 week & 0.32 & 0.91 & 0.93  & 500 lines of code\\
    \framework & 1 hour & 1.53 & 0.91 & 0.90 & 3 MBFs/1 MRE/57 lines\\
    \bottomrule
    \end{tabular}%
    }
  \label{tab:results}%
\end{table}%

@@

For the temporal and number normalization cases, 
we evaluated the techniques against arbitrary text from issues of 
the Lebanese Assafir and Al-Akhbar newspapers
\footnote{available at \url{http://www.assafir.com} and 
\url{http://www.al-akhbar.com}.}. 
For the narrator chain case, we used
Musnad Ahmad, a hadith book, for evaluation. 
For the genealogical tree extraction we used 
an extract from the Genesis biblical text.

@@

\framework required reasonably more runtime than the task specific 
implementations and reported acceptable and 
slightly less precision metrics with around
the same recall.
%a runtime of 7.29, 1.53, 3.14, and 2.28 seconds for 
%the narrator chain, temporal, genealogy, and number normalization tasks, respectively. 
%However, ANGE, ATEEMA, GENTREE, and NUMNORM required 1.79, 2.53, 0.74, and 0.32 seconds, respectively. 
%\framework required a runtime of 7.29, 1.53, 3.14, and 2.28 seconds for 
%the narrator chain, temporal, genealogy, and number normalization tasks, respectively. 
%However, ANGE, ATEEMA, GENTREE, and NUMNORM required 1.79, 2.53, 0.74, and 0.32 seconds, respectively. 
%In narrator chain, \framework scored 0.99\% recall and 0.93\% precision while ANGE scored 0.99\% recall and 0.99\% precision. 
%In temporal entity, \framework scored 0.91\% recall and 0.81\% precision while ATEEMA scored 0.88\% recall and 0.89\% precision. 
%In genealogy, \framework scored 0.84\% recall and 0.93\% precision while GENTREE scored 0.96\% recall and 0.98\% precision. 
%In number normalization, \framework scored 0.91\% recall and 0.90\% precision while NUMNORM scored 0.91\% recall and 0.93\% precision.

@@

\subsection{Narrator chain case study}
A narrator chain is a sequence of narrators referencing each other. 
%A sample narrator chain is shown in Table~\ref{tab:nchain}. 
The chain includes proper nouns, paternal entities, and referencing entities. 
ANGE uses Arabic morphological analysis, finite state machines, and graph transformations 
to extract named entities and relations including narrator chains~\cite{ZaMaFlairs2012HadithBio}.

\transfalse
MBF \cci{PN} checks the abstract category {\tt Name of Person}. 
MBF \cci{FAM} denotes ``family connector'' and checks the stem gloss ``son''. 
MBF \cci{TOLD} denotes referencing between narrators and checks the disjunction of 
the stems \RL{.hd_t}(spoke to), \RL{`n}(about), \RL{sm`}(heard), \RL{'_hbr}(told), and \RL{'nb-'}(inform). 
MBF \cci{MEAN} checks the stem \RL{`ny}(mean). 
MBFs \cci{BLESS}, \cci{GOD}, \cci{UPONHIM}, and \cci{GREET} check the 
stems \RL{.sll_A}, \RL{Al-ll_ah}, \RL{`ly}, and \RL{sllm}, respectively. 
\transtrue

Table~\ref{tab:nchain} presents the defined MREs. 
MRE {\em name} is one or more \cci{PN} tags optionally followed 
with a \cci{MEAN} tag. 
MRE \cci{nar} denotes narrator which is a complex Arabic name
composed as a sequence of Arabic names (\cci{name}) 
connected with family indicators (\cci{FAM}). 
The \cci{NONE} tags in \cci{nar} allow for unexpected words 
that can occur between names. 
%a complex Arabic name that is constructed
%a \cci{name} tag followed by zero or more 
%sequences of \cci{FAM} tag followed by \cci{name} tag 
%with up to three \cci{NONE} tags. 
MRE \cci{pbuh} denotes a praise phrase often associated with 
the end of a hadith (peace be upon him), 
and is the satisfied by the sequence of
\cci{BLESS}, \cci{GOD}, \cci{UPONHIM}, and \cci{GREET} tags. 
MRE \cci{nchain} denotes narrator chain, 
and is a sequence of narrators (\cci{nar})
separated with \cci{TOLD} tags, and optionally followed
by a \cci{pbuh} tag. 
%nal sequence of up to eight \cci{PN}, \cci{FAM}, or 
%\cci{NONE} tags followed by a \cci{pbuh} tag.

The first row in Table~\ref{tab:nchain} is an example narrator chain,
the second is the transliteration, the third 
shows the MBF tags. Rows 4, 5, and 6 show the 
matches for \cci{name}, \cci{nar}, and \cci{nchain},
respectively.
%
\framework assigns the symbols $s_1$ and $s_2$ for the 
MRE sub-expressions \cci{TOLD} and \cci{nar}, respectively. 
We define the relation $\langle s_2,s_2',s_1\rangle$ 
to relate sequences of narrators with edges labelled by the tags of \cci{TOLD} where 
$s_2'$ denotes the next match of \cci{nar} in the one or more MRE subexpression.
%The narrators in the example shown in Table~\ref{tab:nchain} are \RL{qtybT bn s`yd}, \RL{jryr}, and \RL{`mArT bn Alq`qA`}. 
%\transfalse
%The edges relating the entities are labeled by the word set \{\RL{.hdd_tnA}, \RL{.hdd_tnA}, \RL{`n}\} which contains all the %matches of the MBF \cci{TOLD}.
%\transtrue

Table~\ref{tab:mbfer} shows that \framework detected almost all the MBF matches 
with 99\% recall and 85\% precision and 
extracted user-defined relations with 98\% recall and 99\% precision.

@@

\vspace{-1em}
\subsection{Number normalization}
\label{sec:sec:number}

We implemented a number normalization extractor using \framework and 
compared it with {\em NUMNORM}, a 
C++ implementation for number normalization. 
First, we defined the MBFs \cci{DT}, \cci{H}, and \cci{TMB}
to denote (1) digits and tens, (2) hundreds, and (3) 
thousands, millions, and billions, respectively.
The \cci{num} MRE 
\cci{(DT|TMB|H)+} is one or more \cci{DT}, \cci{TMB}, or \cci{H} tags. 
\framework assigns the symbols $s_1$, $s_2$, and $s_3$ 
for the sub-expressions \cci{DT}, \cci{TMB}, and \cci{H}, respectively. 
%The actions associated with the sub-expressions \cci{DT}, \cci{TMB}, and \cci{H} are presented in the appendix.
Figure~\ref{fig:numnormalgo} shows the actions associated with the \cci{DT}, \cci{TMB}, and \cci{H} subexpressions that cumulatively compute the numeric value of the numeric expression match.
The actions use \framework API to access features of the matches such 
as text (\cci{\$s1.text}) and numeric 
value (\cci{\$s1.number}) of literal numbers such as numbers from one to ten.

Table~\ref{tab:mbfer} shows high accuracy in MBF tagging with 99\% recall and 99\% precision, and
high accuracy in relation extraction with 97\% recall and 95\% precision.

\vspace{-1em}
\subsection{Discussion}
\label{subsec:discuss}
\input{discussion}

>>>>

% Table generated by Excel2LaTeX from sheet 'Sheet1'
\begin{table}[tb!]
  \centering
  \caption{\framework compared to task specific applications.}
  \label{tab:results}%
  \resizebox{\columnwidth}{!}{
    \begin{tabular}{l|l|l|l|ll|l}
     \toprule
      \multirow{2}{*}{Task} & Size & Development& Run & \multicolumn{2}{c|}{Accuracy} & \multirow{2}{*}{Ease of Composition}\\
      & (words) & time & time(s) & Recall & \multicolumn{1}{c|}{Precision} & \\
    \midrule
    %ANGE~\cite{ZaMaFlairs2012HadithBio}
      ANGE & 1,865  
    & 2 months & 1.79 & 0.99 & 0.99 & 3000+ lines of code\\
      \framework &  & 3 hours & 7.24 & 0.99 & 0.93  & 8 MBFs and 4 MREs\\
    \midrule
    %ATEEMA~\cite{ZaMa2012IJCLATime}
      ATEEMA & 1,677 & 1.5 months & 2.53 & 0.88  & 0.89  & 1000+ lines of code  \\
      \framework & & 3 hours & 3.14 & 0.91  & 0.81  & 3 MBFs and 2 MREs\\
    \midrule
    %Genealogy tree~\cite{ZaMaHaCicling2012Entity} 
      Genealogy tree & 1,227 
    & 3 weeks & 0.74 & 0.96 & 0.98 & 3000+ lines of code\\
      \framework &  & 4 hours & 2.28 & 0.84 & 0.93  & 3 MBFs and 3 MREs\\
    \midrule
      NUMNORM &  1,399 & 1 week & 0.32 & 0.91 & 0.93  & 500 lines of code\\
      \framework &  & 1 hour & 1.53 & 0.91 & 0.90 & 3 MBFs/1 MRE/57 lines\\
    \bottomrule
    \end{tabular}%
    }
\vspace{-1em}
\end{table}%

In this section we compare \framework implementations of 
the narrator chain, temporal entity, and genealogy entity extraction 
tasks to the task specific techniques proposed to solve them
in ANGE~\cite{ZaMaFlairs2012HadithBio}, 
ATEEMA~\cite{ZaMa2012IJCLATime},  and
GENTREE~\cite{ZaMaHaCicling2012Entity}, respectively. 
We also compare a \framework number normalization task to 
a task specific implementation. 
We evaluated ANGE with
Musnad Ahmad, a hadith book, where we constructed an annotated 
golden reference containing 1,865 words.
We evaluated ATEEMA 
with articles from issues of  the Lebanese 
Al-Akhbar newspaper where we constructed 
an annotated golden reference containing 1,677 words. 
For the genealogical tree extraction we used 
an extract from the Genesis biblical text with 1,227 words.
Finally, we used an annotated article from the Lebanese Assafir
newspaper with 1,399 words to evaluate the NUMNORM case study
\footnote{available at \url{http://www.assafir.com} and 
\url{http://www.al-akhbar.com}.}. 
%
In the online appendix%
~\footnote{available at ~\url{-ommited-for-anonymity-}}%http://webfea.fea.aub.edu.lb/fadi/pdfs/merfappendix.pdf
, we report on eight additional \framework case studies.
\major{
Manual annotators inspected the outcome and provided corrections where tools 
made mistakes.
The corrections form the manual gold annotation that we compared against.
}

Table~\ref{tab:results} reports the development time,
extraction runtime, recall and precision 
of the output MRE tags, 
the size of the task in lines of code or in number of \framework rules, 
for both the standalone task specific and the \framework implementations.
\major{
The development time measures the time required for 
developing the case study.
For instance, ANGE~\cite{ZaMaFlairs2012HadithBio} required two 
months of development by a research assistant with six and 14 hours of 
course work and teaching duties, respectively.
}

\major{
Table~\ref{tab:results} provides runtime results of \framework 
compared to the task specific implementations while running 
MBF and MRE simulations jointly.
This is a rough estimate of the complexity of the \framework simulator. 
%
The complexity of the MBF simulation is the total number of morphological 
solutions for all the words multiplied by the number of user-defined MBFs.
We do not provide a limit on the number of user defined formulae.
In practice, we did not encounter more than ten formulae per case study.
As for the complexity of MRE simulation, converting the rules into 
non-deterministic finite state machines (NDFSM) is done once. 
Simulating an NDFSM over the MBF tags is potentially exponential. 
In practice all our case studies terminated within a predetermined 
time bound of less than 30 minutes. 
\framework required reasonably more runtime than the task specific 
implementations and reported acceptable and 
slightly less precision metrics with around
the same recall.
}

%Table~\ref{tab:mbfer} reports the accuracy of the output MBF tags and the user-defined relation 
%construction in \framework for each task.

@@

@@

%a runtime of 7.29, 1.53, 3.14, and 2.28 seconds for 
%the narrator chain, temporal, genealogy, and number normalization tasks, respectively. 
%However, ANGE, ATEEMA, GENTREE, and NUMNORM required 1.79, 2.53, 0.74, and 0.32 seconds, respectively. 
%\framework required a runtime of 7.29, 1.53, 3.14, and 2.28 seconds for 
%the narrator chain, temporal, genealogy, and number normalization tasks, respectively. 
%However, ANGE, ATEEMA, GENTREE, and NUMNORM required 1.79, 2.53, 0.74, and 0.32 seconds, respectively. 
%In narrator chain, \framework scored 0.99\% recall and 0.93\% precision while ANGE scored 0.99\% recall and 0.99\% precision. 
%In temporal entity, \framework scored 0.91\% recall and 0.81\% precision while ATEEMA scored 0.88\% recall and 0.89\% precision. 
%In genealogy, \framework scored 0.84\% recall and 0.93\% precision while GENTREE scored 0.96\% recall and 0.98\% precision. 
%In number normalization, \framework scored 0.91\% recall and 0.90\% precision while NUMNORM scored 0.91\% recall and 0.93\% precision.

@@

\subsection{Narrator chain case study}
A narrator chain is a sequence of narrators referencing each other. 
%A sample narrator chain is shown in Table~\ref{tab:nchain}. 
The chain includes proper nouns, paternal entities, and referencing entities. 
ANGE uses Arabic morphological analysis, finite state machines, and graph transformations 
to extract entities and relations including narrator chains~\cite{ZaMaFlairs2012HadithBio}.

\transfalse
Table~\ref{tab:nchain} presents the MREs for the narrator chain case study. 
MBF \cci{PN} checks the abstract category {\tt Name of Person}. 
MBF \cci{FAM} denotes ``family connector'' and checks the stem gloss ``son''. 
MBF \cci{TOLD} denotes referencing between narrators and checks the disjunction of 
the stems \RL{.hd_t}(spoke to), \RL{`n}(about), \RL{sm`}(heard), \RL{'_hbr}(told), and \RL{'nb-'}(inform). 
MBF \cci{MEAN} checks the stem \RL{`ny}(mean). 
MBFs \cci{BLESS}, \cci{GOD}, \cci{UPONHIM}, and \cci{GREET} check the 
stems \RL{.sll_A}, \RL{Al-ll_ah}, \RL{`ly}, and \RL{sllm}, respectively. 
\transtrue

MRE {\em name} is one or more \cci{PN} tags optionally followed 
with a \cci{MEAN} tag. 
MRE \cci{nar} denotes narrator which is a complex Arabic name
composed as a sequence of Arabic names (\cci{name}) 
connected with family indicators (\cci{FAM}). 
The \cci{NONE} tags in \cci{nar} allow for unexpected words 
that can occur between names. 
%a complex Arabic name that is constructed
%a \cci{name} tag followed by zero or more 
%sequences of \cci{FAM} tag followed by \cci{name} tag 
%with up to three \cci{NONE} tags. 
MRE \cci{pbuh} denotes a praise phrase often associated with 
the end of a hadith (peace be upon him), 
and is the satisfied by the sequence of
\cci{BLESS}, \cci{GOD}, \cci{UPONHIM}, and \cci{GREET} tags. 
MRE \cci{nchain} denotes narrator chain, 
and is a sequence of narrators (\cci{nar})
separated with \cci{TOLD} tags, and optionally followed
by a \cci{pbuh} tag. 
%nal sequence of up to eight \cci{PN}, \cci{FAM}, or 
%\cci{NONE} tags followed by a \cci{pbuh} tag.

The first row in Table~\ref{tab:nchain} is an example narrator chain,
the second is the transliteration, the third 
shows the MBF tags. Rows 4, 5, and 6 show the 
matches for \cci{name}, \cci{nar}, and \cci{nchain},
respectively.
%
\framework assigns the symbols $s_1$ and $s_2$ for the 
MRE sub-expressions \cci{TOLD} and \cci{nar}, respectively. 
We define the relation $\langle s_2,s_2',s_1\rangle$ 
to relate sequences of narrators with edges labelled by the tags of \cci{TOLD} where 
$s_2'$ denotes the next match of \cci{nar} in the one or more MRE subexpression.
%The narrators in the example shown in Table~\ref{tab:nchain} are \RL{qtybT bn s`yd}, \RL{jryr}, and \RL{`mArT bn Alq`qA`}. 
%\transfalse
%The edges relating the entities are labeled by the word set \{\RL{.hdd_tnA}, \RL{.hdd_tnA}, \RL{`n}\} which contains all the %matches of the MBF \cci{TOLD}.
%\transtrue
%
Table~\ref{tab:mbfer} shows that \framework detected almost all the MBF matches 
with 99\% recall and 85\% precision and 
extracted user-defined relations with 98\% recall and 99\% precision.

@@

\vspace{-1em}
\subsection{Number normalization}
\label{subsec:numnorm}

We implemented a number normalization extractor using \framework and 
compared it with {\em NUMNORM}, a 
C++ implementation for number normalization. 
First, we defined the MBFs \cci{DT}, \cci{H}, and \cci{TMB}
to denote (1) digits and tens, (2) hundreds, and (3) 
thousands, millions, and billions, respectively.
The \cci{num} MRE 
\cci{(DT|TMB|H)+} is one or more \cci{DT}, \cci{TMB}, or \cci{H} tags. 
\framework assigns the symbols $s_1$, $s_2$, and $s_3$ 
for the sub-expressions \cci{DT}, \cci{TMB}, and \cci{H}, respectively. 
%The actions associated with the sub-expressions \cci{DT}, \cci{TMB}, and \cci{H} are presented in the appendix.
\major{
Figure~\ref{fig:numnormalgo} shows the actions associated with the \cci{DT}, \cci{TMB}, and \cci{H} subexpressions that cumulatively compute the numeric value of the numeric expression match.
The actions use \framework API to access features of the matches such 
as text (\cci{\$s1.text}) and numeric 
value (\cci{\$s1.number}) of literal numbers such as numbers from one to ten.
}
%
Table~\ref{tab:mbfer} shows high accuracy in MBF tagging with 99\% recall and 99\% precision, and
high accuracy in relation extraction with 97\% recall and 95\% precision.

\vspace{-1em}
\subsection{Discussion}
\label{subsec:discuss}
\label{sec:discuss}
\input{discussion}
